---
title: "Titanic"
author: "mason-YHY"
date: "2020/7/3"
output: html_document
---

## 载入数据包
```{r, message = FALSE}
library('ggplot2') # visualization
library('ggthemes') # visualization
library('scales') # visualization
library('dplyr') # data manipulation
library('mice') # imputation
library('randomForest') # classification algorithm
library('reshape2') # data reshape
```

## 第一步：读入并查看数据
```{r, message=FALSE, warning=FALSE}
train <- read.csv('train.csv', stringsAsFactors = F, na.strings = c("NA",""))
test  <- read.csv('test.csv', stringsAsFactors = F, na.strings = c("NA",""))
full  <- bind_rows(train, test) # bind training & test data

# check data
str(full)

# dataset dimensions
dim(full)

# Unique values per column
lapply(full, function(x) length(unique(x))) 
```
**数据内容如下：**
Variable Name | Description
--------------|-------------
Survived      | Survived (1) or died (0)
Pclass        | Passenger's class
Name          | Passenger's name
Sex           | Passenger's sex
Age           | Passenger's age
SibSp         | Number of siblings/spouses aboard（兄弟姐妹、配偶的数量）
Parch         | Number of parents/children aboard（父母、子女的数量）
Ticket        | Ticket number
Fare          | Fare（票价）
Cabin         | Cabin（船舱号码）
Embarked      | Port of embarkation（出发地点）


## 第二步：处理缺失值
### 查看缺失值
```{r}
#summary missing value and visualization
#missing_values <- full %>% summarize_all(funs(sum(is.na(.))/n())) 
#missing_values <- melt(missing_values, variable.name="feature", value.name="missing_pct")
#missing_values <- missing_values[missing_values$missing_pct!=0,]
#ggplot(data =  missing_values, aes(x=reorder(feature,-missing_pct),y=missing_pct)) +
  #geom_bar(stat="identity",fill="red") + coord_flip() + theme_few()
md.pattern(full)
``` 
**结果显示缺失值的变量是Far、Embarked、Age、Survived、Cabin，其中Survived是需要预测的值**

### 补全Fare缺失值

```{r}
full[rownames(full)[is.na(full$Fare)],]
```
因此要补全1044号乘客的票价，票价可能与登船地点和乘客等级有关，因此绘制登船地点为“S”和乘客等级3的票价密度分布图，查看1044号乘客的票价最有可能是多少
```{r
ggplot(full[full$Pclass == '3' & full$Embarked == 'S', ], 
  aes(x = Fare)) +
  geom_density(fill = '#99d6ff', alpha=0.4) + 
  geom_vline(aes(xintercept=median(Fare, na.rm=T)),
    colour='red', linetype='dashed', lwd=1) +
  scale_x_continuous(labels=dollar_format()) +
  theme_few()

# Replace missing fare value with median fare for class/embarkment
full$Fare[1044] <- median(full[full$Pclass == '3' & full$Embarked == 'S', ]$Fare, na.rm = TRUE)
```
**结果显示1044号的票价最有可能在中位数水平**

### 补全Embarked缺失值

```{r}
full[rownames(full)[is.na(full$Embarked)],]
```
因此要补全62号、830号两位获救女士的Embarked信息，登船地点可能与票价有关，统计票价和乘客等级与登船地点的数据绘制盒须图，查看票价80的等级1的乘客最有可能在哪里登船

```{r}
# Get rid of our missing passenger IDs
embark_fare <- full %>%
  filter(PassengerId != 62 & PassengerId != 830)

# Use ggplot2 to visualize embarkment, passenger class, & median fare
ggplot(embark_fare, aes(x = Embarked, y = Fare, fill = factor(Pclass))) +
  geom_boxplot() +
  geom_hline(aes(yintercept=80), 
    colour='red', linetype='dashed', lwd=2) +
  scale_y_continuous(labels=dollar_format()) +
  theme_few()
  
  
full$Embarked[c(62, 830)] <- 'C'
```
**根据结果显示，两位乘客最有可能在“C”登船**


## 合并称呼
```{r}
# Grab title from passenger names
full$Title <- gsub('(.*, )|(\\..*)', '', full$Name)

# Show title counts by sex
table(full$Sex, full$Title)

# Titles with very low cell counts to be combined to "rare" level
rare_title <- c('Dona', 'Lady', 'the Countess','Capt', 'Col', 'Don', 
                'Dr', 'Major', 'Rev', 'Sir', 'Jonkheer')

# Also reassign mlle, ms, and mme accordingly
full$Title[full$Title == 'Mlle']        <- 'Miss' 
full$Title[full$Title == 'Ms']          <- 'Miss'
full$Title[full$Title == 'Mme']         <- 'Mrs' 
full$Title[full$Title %in% rare_title]  <- 'Rare Title'

# Show title counts by sex again
table(full$Sex, full$Title)

full$Surname <- sapply(full$Name,  
                      function(x) strsplit(x, split = '[,.]')[[1]][1])

cat(paste('We have <b>', nlevels(factor(full$Surname)), 
'</b> unique surnames. I would be interested to infer ethnicity based on surname --- another time.'))
```

## 统计乘客在船上的家庭总人数, 分析不同规模家庭之间的生存情况
```{r}
# Create a family size variable including the passenger themselves
full$Fsize <- full$SibSp + full$Parch + 1

# Create a family variable 
full$Family <- paste(full$Surname, full$Fsize, sep='_')

# Use ggplot2 to visualize the relationship between family size & survival
ggplot(full[1:891,], aes(x = Fsize, fill = factor(Survived))) +
  geom_bar(stat='count', position='dodge') +
  scale_x_continuous(breaks=c(1:11)) +
  labs(x = 'Family Size') +
  theme_few()
  
  # Discretize family size
full$FsizeD[full$Fsize == 1] <- 'singleton'
full$FsizeD[full$Fsize < 5 & full$Fsize > 1] <- 'small'
full$FsizeD[full$Fsize > 4] <- 'large'

# Show family size by survival using a mosaic plot
mosaicplot(table(full$FsizeD, full$Survived), main='Family Size by Survival', shade=TRUE)
```
家庭规模在2-4人的时候生存较好，其它情况生存较差

## 根据乘客船票信息挖掘乘客在轮船的夹板位置
```{r}
# Create a Deck variable. Get passenger deck A - F:
full$Deck<-factor(sapply(full$Cabin, function(x) strsplit(x, NULL)[[1]][1]))
```
## 根据船票价格、乘客等级、登船位置进行补全数据
### 补全62、830号乘客的登船信息
```{r}
# Get rid of our missing passenger IDs
embark_fare <- full %>%
  filter(PassengerId != 62 & PassengerId != 830)

# Use ggplot2 to visualize embarkment, passenger class, & median fare
ggplot(embark_fare, aes(x = Embarked, y = Fare, fill = factor(Pclass))) +
  geom_boxplot() +
  geom_hline(aes(yintercept=80), 
    colour='red', linetype='dashed', lwd=2) +
  scale_y_continuous(labels=dollar_format()) +
  theme_few()
  
  
full$Embarked[c(62, 830)] <- 'C'
```
### 补全1044号乘客票价信息
```{r}
full[1044, ]
ggplot(full[full$Pclass == '3' & full$Embarked == 'S', ], 
  aes(x = Fare)) +
  geom_density(fill = '#99d6ff', alpha=0.4) + 
  geom_vline(aes(xintercept=median(Fare, na.rm=T)),
    colour='red', linetype='dashed', lwd=1) +
  scale_x_continuous(labels=dollar_format()) +
  theme_few()

# Replace missing fare value with median fare for class/embarkment
full$Fare[1044] <- median(full[full$Pclass == '3' & full$Embarked == 'S', ]$Fare, na.rm = TRUE)
```
## 利用mice包对乘客年龄进行补全
```{r}
# Make variables factors into factors
factor_vars <- c('PassengerId','Pclass','Sex','Embarked',
                 'Title','Surname','Family','FsizeD')

full[factor_vars] <- lapply(full[factor_vars], function(x) as.factor(x))

# Set a random seed
set.seed(129)

# Perform mice imputation, excluding certain less-than-useful variables:
mice_mod <- mice(full[, !names(full) %in% c('PassengerId','Name','Ticket','Cabin','Family','Surname','Survived')], method='rf') 
# Save the complete output 
mice_output <- complete(mice_mod)

# Plot age distributions
par(mfrow=c(1,2))
hist(full$Age, freq=F, main='Age: Original Data', 
  col='darkgreen', ylim=c(0,0.04))
hist(mice_output$Age, freq=F, main='Age: MICE Output', 
  col='lightgreen', ylim=c(0,0.04))

# Replace Age variable from the mice model.
full$Age <- mice_output$Age
```
## 根据年龄等相关信息进一步挖掘乘客是否为儿童、母亲
 A child will simply be someone under 18 years of age and a mother is a passenger who is 1) female, 2) is over 18,   
 3) has more than 0 children (no kidding!), and 4) does not have the title ‘Miss’.
```{r}
# First we'll look at the relationship between age & survival
ggplot(full[1:891,], aes(Age, fill = factor(Survived))) + 
  geom_histogram() + 
  # I include Sex since we know (a priori) it's a significant predictor
  facet_grid(.~Sex) + 
  theme_few()
  
# Create the column child, and indicate whether child or adult
full$Child[full$Age < 18] <- 'Child'
full$Child[full$Age >= 18] <- 'Adult'

# Show counts
table(full$Child, full$Survived)

# Adding Mother variable
full$Mother <- 'Not Mother'
full$Mother[full$Sex == 'female' & full$Parch > 0 & full$Age > 18 & full$Title != 'Miss'] <- 'Mother'

# Show counts
table(full$Mother, full$Survived)

# Finish by factorizing our two new factor variables
full$Child  <- factor(full$Child)
full$Mother <- factor(full$Mother)

md.pattern(full)
```
## 利用随机森林进行预测
```{r}
# Split the data back into a train set and a test set
train <- full[1:891,]
test <- full[892:1309,]

# Set a random seed
set.seed(754)

# Build the model (note: not all possible variables are used)
rf_model <- randomForest(factor(Survived) ~ Pclass + Sex + Age + SibSp + Parch + 
                                            Fare + Embarked + Title + 
                                            FsizeD + Child + Mother,
                                            data = train)

# Show model error
plot(rf_model, ylim=c(0,0.36))
legend('topright', colnames(rf_model$err.rate), col=1:3, fill=1:3)

# Get importance
importance    <- importance(rf_model)
varImportance <- data.frame(Variables = row.names(importance), 
                            Importance = round(importance[ ,'MeanDecreaseGini'],2))

# Create a rank variable based on importance
rankImportance <- varImportance %>%
  mutate(Rank = paste0('#',dense_rank(desc(Importance))))

# Use ggplot2 to visualize the relative importance of variables
ggplot(rankImportance, aes(x = reorder(Variables, Importance), 
    y = Importance, fill = Importance)) +
  geom_bar(stat='identity') + 
  geom_text(aes(x = Variables, y = 0.5, label = Rank),
    hjust=0, vjust=0.55, size = 4, colour = 'red') +
  labs(x = 'Variables') +
  coord_flip() + 
  theme_few()
  
# Predict using the test set
prediction <- predict(rf_model, test)

# Save the solution to a dataframe with two columns: PassengerId and Survived (prediction)
solution <- data.frame(PassengerID = test$PassengerId, Survived = prediction)

# Write the solution to file
write.csv(solution, file = 'rf_mod_Solution.csv', row.names = F)
```

## 载入数据包